```{r}
library(Seurat)
library(Matrix)
library(dplyr)
library(cowplot)
require(data.table)
library(harmony)
library("ggplot2")
library(pals)
library(ggpubr)

options(bitmapType='cairo') 
prefix = "Seurat.MetsGBM.CombinedAnalysis.Apr2022.Myeloid"
infile = "Seurat.MetsGBM.CombinedAnalysis.Apr2022.MinD0.1.sprd5.NN20.seed122.ep1000.rds"
id_file = "Seurat.MetsGBM.CombinedAnalysis.Apr2022.MinD0.1.sprd5.NN20.seed122.ep1000.Myeloid.ID.txt"

```

```{r}
ids <- as.character(read.table(id_file, header = FALSE)[,1])
data.all=readRDS(file=infile)
data.combined = subset(data.all, cells = ids)
rm(data.all)

data.combined@meta.data = data.combined@meta.data[!colnames(data.combined@meta.data) %in% c("integrated_snn_res.0.3","integrated_snn_res.0.5","integrated_snn_res.0.7","seurat_clusters")]

ob.list<-SplitObject(data.combined,split.by= "condition")

ob.list.gbm.rec <- SplitObject(ob.list$GBM.rec,split.by= "ID")
gbm.merge <- merge(
  x = ob.list.gbm.rec[[1]],
  y = unlist(ob.list.gbm.rec[2:length(ob.list.gbm.rec)])
)
gbm.merge@assays$RNA@var.features <- SelectIntegrationFeatures(ob.list.gbm.rec)


ob.list.gbm.pembro <- SplitObject(ob.list$GBM.pembro,split.by= "ID")
gbm.pembro.merge <- merge(
  x = ob.list.gbm.pembro[[1]],
  y = unlist(ob.list.gbm.pembro[2:length(ob.list.gbm.pembro)])
)
gbm.pembro.merge@assays$RNA@var.features <- SelectIntegrationFeatures(ob.list.gbm.pembro)


ob.list.met <- SplitObject(ob.list$met,split.by= "ID")
met.merge <- merge(
  x = ob.list.met[[1]],
  y = unlist(ob.list.met[2:length(ob.list.met)])
)
met.merge@assays$RNA@var.features <- SelectIntegrationFeatures(ob.list.met)


ob.list.met.pembro <- SplitObject(ob.list$met.pembro,split.by= "ID")
met.pembro.merge <- merge(
  x = ob.list.met.pembro[[1]],
  y = unlist(ob.list.met.pembro[2:length(ob.list.met.pembro)])
)
met.pembro.merge@assays$RNA@var.features <- SelectIntegrationFeatures(ob.list.met.pembro)

ob.list = list(gbm.merge, gbm.pembro.merge, met.merge, met.pembro.merge)

rm(ob.list.gbm.rec)
rm(ob.list.gbm.pembro)
rm(ob.list.met)
rm(ob.list.met.pembro)
rm(data.combined)

gc()

```

#Run integration, scaling and PCA
```{r}
k.filterMax <- min(200, min(sapply(ob.list, ncol)))
anchors <- FindIntegrationAnchors(object.list = ob.list, k.filter = k.filterMax, anchor.features = 2000)
data.combined <- IntegrateData(anchorset = anchors, k.weight = k.filterMax)

DefaultAssay(object = data.combined) <- "integrated"

data.combined <- ScaleData(object = data.combined, verbose = TRUE, vars.to.regress = c("percent.mt","percent.ribo","nFeature_RNA","nCount_RNA","S.Score","G2M.Score"))

gc()

data.combined <- RunPCA(object = data.combined, npcs = 30, verbose = TRUE)

rm(anchors)
gc()

```

#Run harmony
```{r}

data.combined <- RunHarmony(data.combined, group.by.vars="Dataset",assay.use = "integrated", plot_convergence = TRUE)
harmony_embeddings <- Embeddings(data.combined, 'harmony')

gc()
```


#Run UMAP for cluster visualization
```{r}
minDist = 0.1
spread1 = 5
NN = 20
seedUse = 122
epoch = 1000

prefix1 = paste(prefix,".MinD",minDist,".sprd",spread1,".NN",NN,".seed",seedUse,".ep",epoch,sep="")

data.combined <- RunUMAP(object = data.combined, reduction = "harmony", dims = 1:30,min.dist = minDist, spread=spread1, n.neighbors=NN, seed.use = seedUse, n.epochs=epoch)

outfile = paste(prefix1,"_UMAP", ".nUMI.png",sep="")
png(filename = outfile,width = 800, height = 800)
FeaturePlot(object = data.combined, features = c("nCount_RNA"), min.cutoff = "q9", pt.size = 2) #umi plot
dev.off()

outfile = paste(prefix1,"_UMAP",".nGene.png",sep="")
png(filename = outfile,width = 800, height = 800)
FeaturePlot(object = data.combined, features = c("nFeature_RNA"), min.cutoff = "q9", pt.size = 2) #umi plot
dev.off()

outfile = paste(prefix1,"_UMAP",".pctRibo.png",sep="")
png(filename = outfile,width = 800, height = 800)
FeaturePlot(object = data.combined, features = c("percent.ribo"), min.cutoff = "q9", pt.size = 2) #umi plot
dev.off()

outfile = paste(prefix1,"_UMAP",".pctMito.png",sep="")
png(filename = outfile,width = 800, height = 800)
FeaturePlot(object = data.combined, features = c("percent.mt"), min.cutoff = "q9", pt.size = 2) #umi plot
dev.off()

data.combined@meta.data$Dataset = factor(data.combined@meta.data$Dataset, levels = 
                                           c("GBM.2021","Mets.2022","Werb.2022","Feldman"))


outfile = paste(prefix1,"_UMAP.dataset.png",sep="")
png(filename = outfile,width = 800, height = 800)
DimPlot(object = data.combined, reduction = "umap", group.by = "Dataset", label = TRUE, label.size = 8, pt.size = 0.5)
dev.off()

outfile = paste(prefix1,"_UMAP.dataset.split.png",sep="")
png(filename = outfile,width = 1600, height = 800)
DimPlot(object = data.combined, reduction = "umap", group.by = "Dataset", split.by = "Dataset", label = FALSE, label.size = 8, pt.size = 0.5)
dev.off()


data.combined@meta.data$condition = factor(data.combined@meta.data$condition, levels = 
                                             c("GBM.rec","GBM.pembro","met","met.pembro"))
color_condition <- c("cyan","blue","pink","red")

outfile = paste(prefix1,"_UMAP.condition.png",sep="")
png(filename = outfile,width = 800, height = 800)
DimPlot(object = data.combined, reduction = "umap", group.by = "condition", label = FALSE, label.size = 8, pt.size = 0.5, cols = color_condition)
dev.off()

outfile = paste(prefix1,"_UMAP.condition.split.png",sep="")
png(filename = outfile,width = 1600, height = 800)
DimPlot(object = data.combined, reduction = "umap", group.by = "condition", split.by = "condition", label = FALSE, label.size = 8, pt.size = 0.5, cols = color_condition)
dev.off()

outfile = paste(prefix1,"_UMAP.WerbCelltype.png",sep="")
png(filename = outfile,width = 800, height = 800)
DimPlot(object = data.combined, reduction = "umap", group.by = "Cell_Type", label = TRUE, label.size = 8, pt.size = 1, na.value = "lightgrey", raster=FALSE)
dev.off()

outfile = paste(prefix1,"_UMAP.Sample.png",sep="")
png(filename = outfile,width = 2000, height = 1200)
DimPlot(object = data.combined, reduction = "umap", group.by = "ID", label = TRUE, label.size = 4, pt.size = 1, na.value = "lightgrey", raster=FALSE)
dev.off()

saveRDS(data.combined, file=paste(prefix1, "rds", sep="."))


```


#Clustering
```{r}
DefaultAssay(object = data.combined) <- "integrated"

data.combined <- FindNeighbors(object = data.combined, reduction = "harmony", dims = 1:30, nn.method = "annoy", annoy.metric = "cosine", force.recalc = TRUE)
data.combined <- FindClusters(data.combined, resolution = 0.7)

#colors = c("red","orange","yellow","green","lightgreen","blue","magenta","cyan","pink","brown","black","darkblue","darkgrey","lightgrey","maroon","lightblue")

outfile = paste(prefix1,"_UMAP", ".Res0.7.png",sep="")
png(filename = outfile,width = 800, height = 800)
DimPlot(object = data.combined, reduction = "umap", group.by = "integrated_snn_res.0.7", label = TRUE, label.size = 8, pt.size = 1)#, cols = colors[1:8])
dev.off()

saveRDS(data.combined, file=paste(prefix1, "rds", sep="."))
#data.combined = readRDS(file=paste(prefix1, ".rds", sep=""))

```


#draw umap of genes of interest ###
```{r}
DefaultAssay(data.combined) <- "RNA"

gene = c("CD14","FCGR3A",
         "S100A8","S100A9",
         "FCN1","CD36",
         "VCAN","CXCR4",
         "CD68","CD163",
         "HLA-DQA1","CLEC9A",
         "IDO1","BATF3",
         "FCER1A","CLEC10A",
         "CD1C","LAMP3",
         "LILRA4","P2RY12", #pDC and MAST cell marker
         "TMEM119","CX3CR1",
         "ADORA3","CCR7",
         "IL1B","CCL2",
         "CXCL8","LGALS3",
         "ISG15","IFIT1",
         "CH25H","ENTPD1",
         "CXCL9","CXCL10",
         "GPNMB","CLEC4E",
         "MRC1","LYVE1",
         "CD3D","MKI67")


gene = c("CD14","FCGR3A",
         "S100A8","S100A9",
         "FCN1","VCAN",
         "CXCL9","CXCL10",
         "ISG15","IFITM1",
         "STAT1","STAT2",
         "HLA-DRB5","HLA-DQA1",
         "HLA-DQB1","CD274",
         "LILRB1","LILRB2",
         "MRC1","LYVE1",
         "VEGFA","IL1B",
         "CCL3","CCL4")

outfile = paste(prefix,"_UMAP.MainCellTypes2.png",sep="")
png(filename = outfile,width = 1200, height = 2000)
FeaturePlot(object = data.combined, features = gene, min.cutoff = "q9",  pt.size = 1, ncol=4, cols = c("lightgrey","blue"))
dev.off()



```


#Find cluster markers
```{r}
DefaultAssay(data.combined) <- "RNA"
Idents(object = data.combined) <- "integrated_snn_res.0.7"
cluster.markers<- FindAllMarkers(object = data.combined, only.pos = TRUE, min.pct = 0.25,logfc.threshold = 0.322,test.use="MAST")
outfile = paste(prefix1,"_ClusterDiffGene.Res0.7", ".csv",sep="")
write.csv(cluster.markers, file=outfile)

saveRDS(data.combined,file=paste(prefix1, ".rds", sep=""))
#data.combined = readRDS(file=paste(prefix1, ".rds", sep=""))

```

#assign cell types ###
```{r}
#assign each cluster cell type
Idents(object = data.combined) <- "integrated_snn_res.0.7"
current.cluster.ids <- c(0:24)
new.cluster.ids <- c("Microglia-ISG-IL10","Complement/RiboHigh","Mphage-IL1B-CCL3-CCL4", #2
                     "Mphage-Angiogenesis","Microglia","Mphage-MRC1-LYVE1",#5
                    "Mphage-Lysosome","Microglia-Phagocytosis","Myeloid.Cycling",#8
                    "cDC2","Monocyte-ISG","Mphage+Tcell",#11
                   "Microglia","MitoHigh","Microglia", #14
                   "Mphage+Tcell","Mphage+Tumor","Mphage+Tumor", #17
                   "Monocyte","Mphage-ISG-MRC1-LILRB2",#19
                   "Mphage+Tumor","MitoHigh","cDC1",#22
                   "Bcell","pDC")
        

Idents(object = data.combined) <- plyr::mapvalues(x = Idents(object = data.combined), from = current.cluster.ids, to = new.cluster.ids)

data.combined@meta.data$celltype <- Idents(object = data.combined)
data.combined@meta.data$celltype <- factor(data.combined@meta.data$celltype,
                                           levels = c("Microglia",
                                                      "Microglia-ISG-IL10",
                                                      "Microglia-Phagocytosis",
                                                      "Complement/RiboHigh",
                                                      "Monocyte",
                                                      "Monocyte-ISG",
                                                      "Mphage-ISG-MRC1-LILRB2",
                                                      "Mphage-Angiogenesis",
                                                      "Mphage-MRC1-LYVE1",
                                                      "Mphage-Lysosome",
                                                      "Mphage-IL1B-CCL3-CCL4",
                                                      "cDC2",
                                                      "cDC1",
                                                      "pDC",
                                                      "Myeloid.Cycling",
                                                      "Mphage+Tcell",
                                                      "Mphage+Tumor",
                                                      "Bcell",
                                                      "MitoHigh"))

#pal.bands(tol(12))
colorsCluster = c(tol(n=12)[3:5],"#999933","#FFCDF3","#CC6677",brewer.paired(12)[7:8],"#FF1493","#C71585","#F0E68C	","#C19A6B","#CD7F32","#6E260E","#FFDAB9","darkgrey","darkgrey","grey","lightgrey")

outfile = paste(prefix1,"_UMAP.Celltype.png",sep="")
png(filename = outfile,width = 1000, height = 800)
DimPlot(object = data.combined, reduction = "umap", group.by = "celltype", label = TRUE, label.size = 5, pt.size = 1, repel=TRUE, cols = colorsCluster, raster=FALSE)
dev.off()

# outfile = paste(prefix1,"_UMAP.Celltype.noLabel.png",sep="")
# png(filename = outfile,width = 1000, height = 800)
outfile = paste(prefix1,"_UMAP.Celltype.noLabel.tiff",sep="")
tiff(file=outfile, width=4000, height=2800, res=600)
DimPlot(object = data.combined, reduction = "umap", group.by = "celltype", label = FALSE, label.size = 8, pt.size = 0.2, cols = alpha(colorsCluster,1))
dev.off()

saveRDS(data.combined,file=paste(prefix1, ".rds", sep=""))
#data.combined = readRDS(file=paste(prefix1, ".rds", sep=""))
```


#Find celltype markers
```{r}
DefaultAssay(data.combined) <- "RNA"
Idents(object = data.combined) <- "celltype"

cluster.markers<- FindAllMarkers(object = data.combined, only.pos = TRUE, min.pct = 0.25,logfc.threshold = 0.322,test.use="MAST")

cluster.markers$DeltaPerLogFC = cluster.markers$avg_log2FC*(cluster.markers$pct.1-cluster.markers$pct.2)
  
outfile = paste(prefix1,"_CelltypeDiffGene", ".csv",sep="")
write.csv(cluster.markers, file=outfile)

```



#draw boxplots of genes of interest ###
```{r}
library(ggplot2)
library(reshape)
library(dplyr)
library(ggpubr)
library(scales)

Celltypes = c("Microglia",
              "Microglia-ISG-IL10",
              "Microglia-Phagocytosis",
              "Complement/RiboHigh",
              "Monocyte",
              "Monocyte-ISG",
              "Mphage-ISG-MRC1-LILRB2",
              "Mphage-Angiogenesis",
              "Mphage-MRC1-LYVE1",
              "Mphage-Lysosome",
              "Mphage-IL1B-CCL3-CCL4",
              "cDC2",
              "cDC1",
              "pDC",
              "Myeloid.Cycling",
              "Mphage+Tcell",
              "Mphage+Tumor",
              "Bcell",
              "MitoHigh")

gene = c("CD14","FCGR3A","CD163","CD68","MKI67")
gene = c("CX3CR1","TMEM119","P2RY12")
gene = c("LRP1","DAB2")
gene = c("CXCL9","CXCL10","ISG15","STAT1","STAT2") 
gene = c("C1QA","C1QB")
gene = c("FCN1","S100A8","S100A9")
gene = c("CLEC9A","BATF3","CLEC10A","FCER1A","CLEC4C","IL3RA")
gene = c("HLA-A","HLA-B","HLA-DRA","HLA-DQA1")
gene = c("CD274","LILRB2","JAML")
gene = c("IL1B","CXCL8","NLRP3","TNF","IL1A","NFKB1")
gene = c("CTSD","CSTB","LYZ")
gene = c("VCAN","VEGFA","LYVE1","MRC1")


SubsetData = subset(data.combined, subset = (celltype %in% Celltypes))

data = SubsetData@assays$RNA@data
expr_gene = data[gene,]
expr_gene = as.data.frame(t(expr_gene))

expr_gene$condition = SubsetData@meta.data$celltype
expr_gene$condition = factor(SubsetData@meta.data$celltype,levels=Celltypes)

colnames(expr_gene) = c(gene,"condition")

colors = colorsCluster
#colors = hue_pal()(21)

ncol = length(gene)
nrow1 = length(gene)/ncol
height1 = nrow1*180
width1 = length(levels(expr_gene$condition))*20*ncol

outfile = paste(prefix,"GeneBoxplot.ByCelltype.MHC.png",sep="_")
png(filename = outfile,width = width1, height = height1)
panels=list()
l=length(colnames(expr_gene))
i=1
for(i in 1:length(gene)){
#for(i in 1:2){
  gene2 = colnames(expr_gene)[i]
  data2 = expr_gene[,c(i,l)]
  colnames(data2)=c("normExpr","condition")
  
  #geom_boxplot(outlier.colour="black", outlier.shape=16,
             #outlier.size=2, notch=FALSE)
 
  y = ggplot(data=data2, aes(x=condition,y=normExpr,fill=condition)) +
       #geom_violin(trim = FALSE)+
       geom_boxplot(outlier.colour="black",notch=TRUE)+
        scale_fill_manual(values=colors)+
      theme(
        axis.text.x = element_blank(),
        axis.text.y = element_text(size=28),
        axis.ticks.x = element_blank(),
        axis.title = element_blank(),
        legend.position = "none",
        plot.title=element_text(face = "italic",size=32))+
      ggtitle(gene2)
    
      panels[[i]]=y
  
}
ggarrange(plotlist=panels,ncol=ncol,nrow=nrow1)
dev.off()


Idents(object = SubsetData) <- "celltype"
AverageExpr <- AverageExpression(SubsetData,
                                 assays = "RNA",
                                 features = gene,
                                 group.by= "ident",
                                 return.seurat = TRUE)

outfile = paste(prefix,"_AverageExprByCelltype.txt",sep="")
write.table(AverageExpr@assays$RNA@scale.data,file=outfile,sep="\t")

genelist = c("CD163","MRC1","LYVE1","VEGFA")

outfile = paste(prefix,"Dotplot.SelectedGene.png",sep=".")
png(outfile, height = 500, width = 800)
DotPlot(SubsetData, features = genelist)
dev.off()


```


#calculate each cell type's fraction in each sample ###
```{r}
## Enumerate per sample
totalNum=c(table(data.combined@meta.data$ID)[[1]])
Group = names(table(data.combined@meta.data$ID))
table_sub = data.frame()
table_sub = subset(data.combined@meta.data, ID==Group[1])
percentages=as.data.frame((table(table_sub$celltype)))$Freq
table1 = percentages

m1 = data.combined@meta.data

for(i in 2:length(Group)){
  id = Group[i]
  n_total = table(data.combined@meta.data$ID)[[i]]
  totalNum=c(totalNum,n_total)
  
  table_sub = data.frame()
  table_sub = subset(data.combined@meta.data, ID==id)
  
  percentages=as.data.frame((table(table_sub$celltype)))$Freq
  table1=as.data.frame(rbind(table1,percentages))
}

table2 = as.data.frame(cbind(table1,Total=totalNum))
rownames(table2)=Group
colnames(table2)=c(levels(data.combined@meta.data$celltype),"Total")
table2$Myeloid = table2$Total-table2$'Bcell'-table2$'MitoHigh'-table2$'Mphage+Tcell'-table2$'Mphage+Tumor'

sampleMap = unique(data.combined@meta.data[,c("ID","condition","CancerType","MainCancerType")])
rownames(sampleMap) = sampleMap$ID
sampleMap = sampleMap[Group,]
table2$condition = sampleMap$condition
table2$CancerType = sampleMap$CancerType
table2$MainCancerType = sampleMap$MainCancerType

outfile = paste(prefix,"_CelltypeByID.txt",sep="")
write.table(table2,file=outfile,sep="\t")

```


## Get DEGs for all/subset of myeloid cells
```{r}
DefaultAssay(data.combined) <- "RNA"

Group = names(table(data.combined@meta.data$ID))
sampleMap = unique(data.combined@meta.data[,c("ID","condition","CancerType")])
rownames(sampleMap) = sampleMap$ID
sampleMap = sampleMap[Group,]

Celltypes = c("Microglia",
              "Microglia-ISG-IL10",
              "Microglia-Phagocytosis",
              "Monocyte",
              "Monocyte-ISG",
              "Mphage-ISG-MRC1-LILRB2",
              "Mphage-Angiogenesis",
              "Mphage-MRC1-LYVE1",
              "Mphage-Lysosome",
              "Mphage-IL1B-CCL3-CCL4",
              "cDC2",
              "cDC1",
              "pDC")#,
              #"Complement/RiboHigh",
              #"Myeloid.Cycling")#,
              #"Mphage+Tcell",
              #"Mphage+Tumor",
              #"Bcell",
              #"MitoHigh")

SubsetData = subset(data.combined, subset = (celltype %in% Celltypes))

#sample the same number of cells from every sample to avoid patient-specific effects
IDlist = c()
nCell = 50
set.seed(123)
for(i in 1:length(Group)){
  ids_sub = c()
  
  subdata = subset(SubsetData,ID==Group[i])
  ids = colnames(subdata)
  
  if(length(ids) < nCell){
     ids_sub = ids
  }else{
     ids_sub = sample(ids, nCell)
  }
  IDlist = c(IDlist,ids_sub)
}


## get DEG in met vs GBM samples ####
conditions = c("GBM.rec","met")
TreatmentSubset = subset(data.combined[,IDlist],condition %in% conditions)
Idents(object = TreatmentSubset) <- "condition"

markers <- FindAllMarkers(TreatmentSubset,min.pct = 0.25, logfc.threshold = 0.322,only.pos = TRUE)
markers <- as.data.frame(markers)
markers$diff = markers$pct.1 - markers$pct.2
markers$DeltaPerMultiLog2FC = markers$diff*markers$avg_log2FC

outfile = paste(prefix,"DEG","BrMvsGBM.NoPembro","NoMicrogliaNoDoubletNoB.tsv",sep=".")
write.table(markers, file=outfile, sep="\t" )


## get DEG in BrM.ICB vs BrM samples ####
conditions = c("met","met.pembro")
TreatmentSubset = subset(data.combined[,IDlist],condition %in% conditions)

Idents(object = TreatmentSubset) <- "condition"
markers <- FindAllMarkers(TreatmentSubset,min.pct = 0.25, logfc.threshold = 0.322,only.pos = TRUE)
markers <- as.data.frame(markers)
markers$diff = markers$pct.1 - markers$pct.2
markers$DeltaPerMultiLog2FC = markers$diff*markers$avg_log2FC

outfile = paste(prefix,"DEG","MetsICBvsMets","NoDoubletNoBNoPlasma.tsv",sep=".")
write.table(markers, file=outfile, sep="\t" )


# Get the delta of the delta (GBM.pembro vs GBM.rec, Mets.pembro vs Mets)
conditions = c("GBM.rec","GBM.pembro")
TreatmentSubset = subset(data.combined[,IDlist],condition %in% conditions)
Idents(object = TreatmentSubset) <- "condition"
markers1 <- FindAllMarkers(TreatmentSubset,min.pct = 0, logfc.threshold = 0,only.pos = FALSE)

conditions = c("met","met.pembro")
TreatmentSubset = subset(data.combined[,IDlist],condition %in% conditions)
Idents(object = TreatmentSubset) <- "condition"
markers2 <- FindAllMarkers(TreatmentSubset,min.pct = 0, logfc.threshold = 0,only.pos = FALSE)

markers1.sub = subset(markers1, cluster == "GBM.pembro")
markers2.sub = subset(markers2, cluster == "met.pembro")

genes = union(unique(markers1.sub$gene),unique(markers2.sub$gene))

DeltaLogFC = data.frame(matrix(ncol = 4, nrow = 0))
colnames(DeltaLogFC) = c("Gene","Delta","GBM.pembro.logFC","Mets.Pembro.logFC")

for(i in 1:length(genes)){
  gene = genes[i]
  list1 = markers1.sub[markers1.sub$gene==gene,]
  list2 = markers2.sub[markers2.sub$gene==gene,]
  
  if(nrow(list1)==1){
    logfc1 = list1$avg_log2FC
  }else{
    logfc1 = 0
  }

  if(nrow(list2)==1){
    logfc2 = list2$avg_log2FC
  }else{
    logfc2 = 0
  }  
  
  delta = logfc2-logfc1
  
  DeltaLogFC[i,1] = gene
  DeltaLogFC[i,2] = delta
  DeltaLogFC[i,3] = logfc1
  DeltaLogFC[i,4] = logfc2

}
rownames(DeltaLogFC) = DeltaLogFC$Gene

outfile = paste(prefix,"MetsPembrovsGBMpembro","NoDoubletNoBNoPlasma","DeltaLogFC","tsv",sep=".")
write.table(DeltaLogFC, file=outfile, sep="\t")


```

## get sample-level averaged gene expression boxplot/violin plot
```{r}
SubsetData = subset(data.combined, subset = (celltype %in% Celltypes))

Group = names(table(SubsetData@meta.data$ID))
sampleMap = unique(SubsetData@meta.data[,c("ID","condition","CancerType","MainCancerType","MainCancerType2")])
rownames(sampleMap) = sampleMap$ID
sampleMap = sampleMap[Group,]

#gene sample-wise gene expression level
Idents(SubsetData) <- "ID"
expr_avg <- AverageExpression(SubsetData,
                              assays = "RNA",
                              group.by = "ident",
                              #add.ident = "condition",
                              slot = "data")
expr_avg = as.data.frame(expr_avg$RNA)
expr_avg = expr_avg[,Group]

Genes = c("LYZ","S100A8","S100A9","G0S2","JAML","MRC1","CXCR4","NLRP3","CH25H",
          "CXCL2","CXCL3","CXCL8","COX7B","COX7C","NDUFB1","CCL3","CD44",
          "ISG15","ISG20","IFI6","IFI30","IFI27L2","IFIT1","STAT1","STAT2","TAP1",
          "IFI27","IFIT3","IFITM1","MX1","MX2","OAS2","OASL","GBP5","GBP4",
          "HLA-A","HLA-B","HLA-C","HLA-DRB5","HLA-DQA2")

expr_gene = expr_avg[unique(Genes),sampleMap$ID]
expr_gene = as.data.frame(t(expr_gene))

expr_gene$CancerType = sampleMap$CancerType
expr_gene$condition = sampleMap$condition
expr_gene$MainCancerType = sampleMap$MainCancerType
expr_gene$MainCancerType2 = sampleMap$MainCancerType2

colnames(expr_gene) = c(unique(Genes),"type","condition","MainType","MainType2")

outfile = paste(prefix,"_GeneExprByID.txt",sep="")
write.table(expr_gene,file=outfile,sep="\t")


```

## plot per-celltype-based gene expression boxplot/violin plot
```{r}
Genes = c("LYZ","S100A8","S100A9","JAML","CXCL2",
          #"CXCL3","CXCL8","COX7B","COX7C","NDUFB1",
          "ISG15","ISG20","IFI27L2","HLA-B","HLA-C")

Genes = c("STAT1","OASL","IFITM1","IFITM2","ISG15","ISG20","IFI27","IFI30","IFI6","GBP5","TAP1","HLA-A","HLA-B","HLA-DRA","HLA-DRB5","HLA-DQA2","HLA-DQB2","CXCR4","NLRP3","CH25H")

Genes = c("CXCR4","CXCL12")

Celltypes = c("Microglia",
              "Microglia-ISG-IL10",
              "Microglia-Phagocytosis",
              "Complement/RiboHigh",
              "Monocyte",
              "Monocyte-ISG",
              "Mphage-ISG-MRC1-LILRB2",
              "Mphage-Angiogenesis",
              "Mphage-MRC1-LYVE1",
              "Mphage-Lysosome",
              "Mphage-IL1B-CCL3-CCL4",
              "cDC2",
              "cDC1",
              "pDC",
              "Myeloid.Cycling")#,
              #"Mphage+Tcell",
              #"Mphage+Tumor",
              #"Bcell",
              #"MitoHigh")

condition_sub = c("met","met.pembro")
#condition_sub = c("GBM.rec","GBM.pembro","met","met.pembro")

#cancertype_sub = c("Melanoma")
cancertype_sub = c("Melanoma","Lung","Gastric","Colorectal","RCC","Ovarian","Endometrial","Uterine","Recurrent","Breast","Adenoid.cystic","Unknown")


SubsetData = subset(data.combined, celltype %in% Celltypes & condition %in% condition_sub & CancerType %in% cancertype_sub)

Group = names(table(SubsetData@meta.data$ID))

#sample the same number of cells from every sample to avoid patient-specific effects
IDlist = c()
nCell = 50
set.seed(123)
for(i in 1:length(Group)){
  ids_sub = c()
  
  subdata = subset(SubsetData,ID==Group[i])
  ids = colnames(subdata)
  
  if(length(ids) < nCell){
     ids_sub = ids
  }else{
     ids_sub = sample(ids, nCell)
  }
  IDlist = c(IDlist,ids_sub)
}

TreatmentSubset = data.combined[,IDlist]

DefaultAssay(TreatmentSubset)<-"RNA"

expr_gene = TreatmentSubset@assays$RNA@data[Genes,]
expr_gene = as.data.frame(t(expr_gene))

expr_gene$type = TreatmentSubset@meta.data$celltype
expr_gene$condition = TreatmentSubset@meta.data$condition
expr_gene$MainType = TreatmentSubset@meta.data$MainCancerType

colnames(expr_gene) = c(Genes,"type","condition","MainType")

expr_gene <- droplevels(expr_gene)
expr_gene$type = factor(expr_gene$type, levels = Celltypes)

colors = c("pink","red")
#colors = c("cyan","blue","pink","red")

ncol = 1  
nrow1 = length(Genes)/ncol
width1 = length(Celltypes)*50*ncol
height1 = nrow1*180

outfile = paste(prefix,"DEG.Boxplot","MetsPembrovsGBMpembro","NoDoubletNoBNoPlasma","DeltaLogFC","CXCL12","png", sep=".") #gene expr boxplot in each cell type
png(filename = outfile,width = width1, height = height1)
panels=list()
l=length(colnames(expr_gene))

for(j in 1:length(Genes)){
#for(i in 1:2){
  gene2 = colnames(expr_gene)[j]
  data2 = expr_gene[,c(j,l-2,l-1,l)]
  colnames(data2)=c("normExpr","type","condition","MainType")
 
  y = ggplot(data=data2, aes(x=type,y=normExpr,fill=condition)) +
       #geom_violin(trim = FALSE)+
      geom_boxplot(outlier.colour=alpha("grey",alpha = 0),notch=TRUE)+
      geom_point(aes(fill = condition), size = 0.5, shape = 21, position = position_jitterdodge())+
      scale_fill_manual(values=rep(colors,length(levels(expr_gene$type))))+
      theme(
        #axis.text.x = element_blank(),
        axis.text.x = element_text(size=8),
        axis.text.y = element_text(size=28),
        axis.ticks.x = element_blank(),
        axis.title = element_blank(),
        legend.position = "none",
        plot.title=element_text(face = "italic", size=36))+
      ggtitle(gene2)
    
      panels[[j]]=y
  
}
ggarrange(plotlist=panels,ncol=ncol,nrow=nrow1)
dev.off()



### pairwise wilcox test ###########################
for(j in 1:length(Genes)){
  gene2 = colnames(expr_gene)[j]
  print(gene2)
  
  for(i in 1:length(Celltypes)){
    data2 = expr_gene[,c(j,l-2,l-1,l)]
    colnames(data2)=c("normExpr","type","condition","MainType")
    data2.sub = data2[data2$type %in% c(Celltypes[i]) & data2$condition %in% c("met","met.pembro"),]
    data2.sub <- droplevels(data2.sub)
    t <- wilcox.test(normExpr ~ condition, data = data2.sub)
    
    print(Celltypes[i])
    print(t$p.value)
  }
}




```



