
```{r}
library(Seurat) 
library(AUCell)
library(GSEABase)
library(Matrix)
library(dplyr)
library(matrixStats)
library(pheatmap)
library(klaR)
library(scales)
library(RColorBrewer)
library(viridis)
library(stats)
library(factoextra)
library(graph4lg)
library(reshape)
library(ggpubr)


options(bitmapType='cairo') #to be able to run png() on severs without X11


infile = "Visium.MetsGBM.CombinedAnalysis.Aug2022.rds"

prefix = "Visium.MetsGBM.CombinedAnalysis.Aug2022.SingleSetRef.AUCell.MetsGBMscRNAseqCelltype.SingleSetRef"

gmtFile = "/home/hugolab/DataRAID/GSVA/GeneSets/Seurat.MetsGBM.CombinedAnalysis.Apr2022.All.CelltypeMarker.gmt"

```


### load visium data and self-defined gene set file and run AUCell##################
```{r}
data.combined=readRDS(file=infile)

exprMatrix = as.matrix(data.combined@assays$Spatial@data)
#exprMatrix = as.matrix(data.combined@assays$Spatial@counts)
#exprMatrix = as.matrix(data.combined@assays$SCT@data)

cells_rankings <- AUCell_buildRankings(exprMatrix,nCores=4)

dev.off()

#Note that the size of the gene-set will also affect the results. With smaller gene-genes 
#(fewer genes), it is more likely to get cells with AUC = 0. While this is the case of the 
#"perfect markers" it is also easier to get it by chance with smal datasets. (i.e. Random gene 
#set with 50 genes in the figure). Bigger gene-sets (100-2k) can be more stable and easier to 
#evaluate, as big random gene sets will approach the normal distibution.
geneSets <- getGmt(gmtFile)

#In order to calculate the AUC, by default only the top 5% of the genes in the ranking are used 
#(i.e. checks whether the genes in the gene-set or signature are within the top 5%).
#The percentage to take into account can be modified with the argument aucMaxRank. 
#For datasets where most cells express many genes (e.g. a filtered dataset), or these have high expression values, 
#it might be good to increase this threshold.
#Check the histogram provided by  AUCell_buildRankings to get an estimation on where this threshold lies within the dataset.
cells_AUC <- AUCell_calcAUC(geneSets, cells_rankings, nCores = 4)

set.seed(123)
cells_assignment <- AUCell_exploreThresholds(cells_AUC, plotHist=TRUE, assign=TRUE) 
dev.off()

outfile = paste(prefix,"RData",sep=".")
save(cells_rankings, cells_AUC, cells_assignment, file=outfile)
#load(outfile)

warningMsg <- sapply(cells_assignment, function(x) x$aucThr$comment)
warningMsg[which(warningMsg!="")]


```

#### create negative control distribution by running AUCell on the cell type-absent spots #####
```{r}

celltype = c("Myeloid","Lymphoid","Bcell")
probs <- c(0.95)

subtype = "Bcell"
gmtFile = "/home/hugolab/DataRAID/GSVA/GeneSets/Seurat.MetsGBM.CombinedAnalysis.Apr2022.CelltypeMarker.Bcell.gmt"

geneSets <- getGmt(gmtFile)

id_file = paste("Visium.MetsGBM.CombinedAnalysis.Aug2022.SingleSetRef", "No", subtype, "ID.txt" ,sep=".")
ids <- as.character(read.table(id_file, header = FALSE)[,1])
  
data.sub = subset(data.combined, cells = ids)
exprMatrix_sub = as.matrix(data.sub@assays$Spatial@data)

cells_rankings_sub <- AUCell_buildRankings(exprMatrix_sub, plotStats=FALSE, nCores=4)
cells_AUC_sub <- AUCell_calcAUC(geneSets, cells_rankings_sub, nCores = 4)
AUC_score = as.matrix(cells_AUC_sub@assays@data$AUC)

set.seed(123)
# outfile = paste(prefix, "NullDist", subtype, "pdf", sep = ".")
# pdf(outfile)
cells_assignment_sub <- AUCell_exploreThresholds(cells_AUC_sub, plotHist=TRUE, assignCells=FALSE)
dev.off()

q <- rowQuantiles(AUC_score, probs = probs)
q = as.data.frame(q)
rownames(q) = names(geneSets)

write.table(q, file=paste(prefix,"SubtypeAbsent.AUCThreshold",subtype,"txt",sep="."), sep="\t")

##########################################################################################
## combine the tables after run threshold for all cell types##############################
AUC_Threshold = as.data.frame(matrix(ncol=1,nrow=0))
for(i in 1:length(celltype)){
  subtype = celltype[i]
  
  file=paste(prefix,"SubtypeAbsent.AUCThreshold",subtype,"txt",sep=".")
  q = read.delim(file, sep = "\t")
  
  AUC_Threshold = rbind(AUC_Threshold, q)
}

write.table(AUC_Threshold, file=paste(prefix,"SubtypeAbsent.AUCThreshold","txt",sep="."), sep="\t")


```

#### Assign new thresholds calculated based on the cell type-absent negative distribution #####
```{r}

outfile = paste(prefix,"RData",sep=".")
load(outfile)

AUC_Threshold = read.delim(paste(prefix,"SubtypeAbsent.AUCThreshold","txt",sep="."),row.names=1,sep = "\t")

for(i in 1:length(rownames(cells_AUC))){
  #i = i+1
  subtype = rownames(cells_AUC)[i]
  subtype
  
  threshold = AUC_Threshold[subtype,]

  newSelectedCells <- names(which(getAUC(cells_AUC)[subtype,]>threshold))
  length(newSelectedCells)
  
  intersect(newSelectedCells,cells_assignment[[subtype]]$assignment)
  
  cells_assignment[[subtype]]$assignment <- newSelectedCells
  #text = paste("AUC>",round(threshold,digits=2),"(",length(newSelectedCells),")",sep="")
  
  subtype2 = gsub("/", ".", subtype)
  
  outfile = paste(prefix, "SubtypeAbsent.AUCThreshold",subtype2, "png",sep=".")
  png(filename = outfile,width = 1000, height = 1000)
  AUCell_plotHist(cells_AUC[subtype,], aucThr=threshold)
  abline(v=threshold, col="red")
  dev.off()
}

outfile = paste(prefix,"RData",sep=".")
save(cells_rankings, cells_AUC, cells_assignment, file=outfile)
#load(outfile)

```

### Extract these cells for all the gene-sets and transform it into a table, do this recursively for multiple gmt files ####
```{r}

infile_list = list("Visium.MetsGBM.CombinedAnalysis.Aug2022.AUCell.MetsGBMscRNAseqCelltype.RData")

assignmentTable_combined = data.frame(matrix(ncol=2, nrow=0))

for(i in 1:length(infile_list)){
  load(infile_list[[i]])
  cellsAssigned <- lapply(cells_assignment, function(x) x$assignment)
  assignmentTable <- reshape2::melt(cellsAssigned, value.name="cell")
  colnames(assignmentTable)[2] <- "geneSet"
  head(assignmentTable)
  
  assignmentTable_combined = rbind(assignmentTable_combined,assignmentTable)
}


#For tumor cell and normal brain, EC, vSMC cells, we do not have subtype gene signatures, so we used a percent cutoff >=20% to define their presence
id_file = "Visium.MetsGBM.CombinedAnalysis.Aug2022.SingleSetRef.With.Tumor.ID.txt"
ids <- as.character(read.table(id_file, header = FALSE)[,1])
assignmentTable = data.frame(cell=ids,geneSet=rep("Tumor",length(ids)))
assignmentTable_combined = rbind(assignmentTable_combined,assignmentTable)

id_file = "Visium.MetsGBM.CombinedAnalysis.Aug2022.SingleSetRef.With.BrainNormal.ID.txt"
ids <- as.character(read.table(id_file, header = FALSE)[,1])
assignmentTable = data.frame(cell=ids,geneSet=rep("BrainNormal",length(ids)))
assignmentTable_combined = rbind(assignmentTable_combined,assignmentTable)

id_file = "Visium.MetsGBM.CombinedAnalysis.Aug2022.SingleSetRef.With.Vascular.ID.txt"
ids <- as.character(read.table(id_file, header = FALSE)[,1])
assignmentTable = data.frame(cell=ids,geneSet=rep("Vascular",length(ids)))
assignmentTable_combined = rbind(assignmentTable_combined,assignmentTable)

assignmentMat <- t(table(assignmentTable_combined[,"geneSet"], assignmentTable_combined[,"cell"])) #Convert into an incidence matrix 

assignmentMat = as.data.frame.matrix(assignmentMat)

assignmentMat$rownames = rownames(assignmentMat)

AUC_Binary <- left_join(data.frame(rownames = rownames(data.combined@meta.data)),
                        assignmentMat,
                        by = "rownames") #reorder assignmentMat based on seurat object cell order

rownames(AUC_Binary) = AUC_Binary$rownames

AUC_Binary = AUC_Binary[,!colnames(AUC_Binary) %in% c("rownames")]

data.combined@assays[["AUC_Binary"]] <- CreateAssayObject(data = t(AUC_Binary))

data.combined@assays$AUC_Binary@key <- "newkey_"

DefaultAssay(data.combined) <- "AUC_Binary"

types = colnames(AUC_Binary)

outfile = paste(prefix,"AUC_Binary","png",sep=".")
png(filename = outfile,width = 1500, height = 10000)
#SpatialFeaturePlot(data.combined, features = types)
SpatialFeaturePlot(data.combined, features = types)
dev.off()


celltype = "CD8.CXCL13.LAG3.Tex"
ID_subset = rownames(AUC_Binary[AUC_Binary[,celltype]==1,])

outfile = paste(prefix,"AUC_Binary",celltype,"png",sep=".")
png(filename = outfile,width = 8000, height = 2000)
SpatialDimPlot(data.combined,cells.highlight=ID_subset)
dev.off()



write.table(AUC_Binary, file=paste(prefix,"AUC_Binary","txt",sep="."), sep="\t")
#AUC_Binary <- read.delim(file=paste(prefix,"AUC_Binary","txt",sep="."),sep="\t",header=T,row.names=1)
```


### calculate the number of each cell type ############
```{r}

Samplelist = unique(data.combined@meta.data$ID)
AUC_Binary <- read.delim(file=paste(prefix,"AUC_Binary","txt",sep="."),sep="\t",header=T,row.names=1)
types = colnames(AUC_Binary)

types_Tcell = c("CD4.CXCL13.CTLA4.Tex","CD4.IL7R.CCR7.CD40LG.Tcm",
                "CD8.CXCL13.LAG3.Tex","CD8.GZMH.GZMA.CD52.Teff",
                "CD8.GZMK.CCL4.CCL5.Teff","CD8.IL7R.CD69.Tm",
                "CD8.ISG.High","CD8.ITGA1.ITGAL.Trm",
                "CD8.TCF7.CD226.Tprog.exh","CD8.VCAM1.IFNG.Tex",
                "CTL.NK","CTL.NK.XCL1.XCL2",
                "IL7R.KLRB1.Tm","Treg")


types_Myeloid = c("cDC1","cDC2",
                  "Complement.RiboHigh","Microglia",
                  "Microglia.ISG.IL10","Microglia.Phagocytosis",
                  "Monocyte","Monocyte.ISG",
                  "Mphage.Angiogenesis","Mphage.IL1B.CCL3.CCL4",
                  "Mphage.ISG.MRC1.LILRB2","Mphage.Lysosome",
                  "Mphage.MRC1.LYVE1","pDC")

type_num <- data.frame()
#colnames(type_num) = types
#rownames(type_num) = Samplelist

for(i in 1:length(Samplelist)){
  sample = Samplelist[i]
  ids = colnames(subset(x=data.combined, subset = ID == sample))
  all_type = c() #list of wells with any t cell subsets  
  
  for(j in 1:length(types_Tcell)){
    type = types_Tcell[j]
    subset = AUC_Binary[rownames(AUC_Binary) %in% ids & !is.na(AUC_Binary[,type]) & AUC_Binary[,type]==1,]
    
    n_subtype = nrow(subset)
    
    all_type = union(all_type,rownames(subset))
    
    type_num[i,type] = n_subtype
  }
  type_num[i,"Tcell_total"] <- length(all_type)
}

for(i in 1:length(Samplelist)){
  sample = Samplelist[i]
  ids = colnames(subset(x=data.combined, subset = ID == sample))
  all_type = c() #list of wells with any t cell subsets  
  
  for(j in 1:length(types_Myeloid)){
    type = types_Myeloid[j]
    subset = AUC_Binary[rownames(AUC_Binary) %in% ids & !is.na(AUC_Binary[,type]) & AUC_Binary[,type]==1,]
    
    n_subtype = nrow(subset)
    
    all_type = union(all_type,rownames(subset))
    
    type_num[i,type] = n_subtype
  }
  type_num[i,"Myeloid_total"] <- length(all_type)
}

rownames(type_num) = Samplelist

write.table(type_num,file=paste(prefix,"CelltypeNum","txt",sep="."), sep="\t")


```


### contruct proximity matrix ######################################################
```{r}

#AUC_Binary = AUC_Binary[,!colnames(AUC_Binary) %in% c("ISGhi.Mphage","Proliferating.Lymphoid","Proliferating.Myeloid", "Tumor.cycling","Adipose_Tissue_LYVE1_DN","LUNG_Lyve1loMHCIIhi","Adipose_Tissue_LYVE1_UP")]
Samplelist = unique(data.combined@meta.data$ID)
Imagelist = Images(object = data.combined)

for(i in 1:length(Samplelist)){
  image = Imagelist[i]
  sample = Samplelist[i]
  
  prox_table <- data.frame(matrix(nrow=0,ncol=3)) #average minimum distance from points to centers
  prox_table2 <- data.frame(matrix(nrow=0,ncol=3)) #average distance between centers and points
  
  colnames(prox_table) = c("point_name", "center_name","prox_mean")
  colnames(prox_table2) = c("point_name", "center_name","prox_mean")


  table_coord <- data.combined[[image]]@coordinates
  table_coord <- table_coord[order(table_coord$row, table_coord$col),]
  minDist = sqrt((table_coord[1,"imagerow"]-table_coord[2,"imagerow"])^2+(table_coord[1,"imagecol"]-table_coord[2,"imagecol"])^2)
  minDist = round(minDist,digits=0)

  coord <- GetTissueCoordinates(data.combined[[image]],scale=NULL) #A numeric data.frame (or matrix) representing the spatial pixel locations. r ownames are barcodes/pixel names, and there should be two columns for 'x' and for 'y'
  AUC_sub = AUC_Binary[rownames(coord),]
  
  table = cbind(AUC_sub,coord)
  
  for(j in 1:ncol(AUC_sub)){
    point_name = colnames(AUC_sub)[j]
    points <- subset(table,table[[j]]==1)[,c("imagerow","imagecol")]
    points$point = rownames(points)
    colnames(points) <- c("imagerow_p","imagecol_p","point")

    for(k in j:ncol(AUC_sub)){
      center_name = colnames(AUC_sub)[k]
      centers <- subset(table,table[[k]]==1)[,c("imagerow","imagecol")]
      centers$center = rownames(centers)
      colnames(centers) <- c("imagerow_ce","imagecol_ce","center")
      
      if(nrow(centers) == 0){next}
      #calculate average minimum distance from points to centers
      result1 <- points %>% group_by(point) %>%  distinct() %>%
        mutate(lst = with(centers, list(sqrt( (imagerow_p-imagerow_ce)^2 + (imagecol_p-imagecol_ce)^2 ) ) ),
               distance=min(unlist(lst)),
               center = centers$center[which.min(unlist(lst))]) #%>%
        #select(-lst)
      
      dist_mean = mean(result1$distance)
      dist_mean = dist_mean/minDist
      prox_mean = 1/(1+dist_mean)
      prox_table <-rbind(prox_table, data.frame(point_name = point_name, center_name=center_name, prox_mean = prox_mean))
      
      #average distance between centers and points
      # result1 <- points %>% group_by(point) %>%  distinct() %>% 
      #   mutate(lst = with(centers, list(sqrt( (imagerow_p-imagerow_ce)^2 + (imagecol_p-imagecol_ce)^2 ) ) ), 
      #          distance=mean(unlist(lst))) 
      
      #dist_mean = mean(result1$distance)
      #dist_mean = dist_mean/minDist
      
      point_row = mean(points$imagerow_p)
      point_col = mean(points$imagecol_p)
      
      center_row = mean(centers$imagerow_ce)
      center_col = mean(centers$imagecol_ce)
      
      dist_mean2 = sqrt( (point_row-center_row)^2 + (point_col-center_col)^2 )/minDist
      prox_mean2 = 1/(1+dist_mean2)

      prox_table2 <-rbind(prox_table2, data.frame(point_name = point_name, center_name=center_name, prox_mean = prox_mean2))

    }
  }
  colnames(prox_table) = c("Point","Center","Proximity_AverageMin")
  pw_mat = df_to_pw_mat(prox_table,from="Point", to="Center", value="Proximity_AverageMin")

  write.table(prox_table,file=paste(prefix,"MinMeanProximity",sample,"txt",sep="."), sep="\t")
  write.table(pw_mat,file=paste(prefix,"MinMeanProximity",sample,"Pairwise","txt",sep="."), sep="\t")

  
  colnames(prox_table2) = c("Point","Center","Proximity_Average")
  pw_mat = df_to_pw_mat(prox_table2,from="Point", to="Center", value="Proximity_Average")

  write.table(prox_table2,file=paste(prefix,"AverageProximity",sample,"txt",sep="."), sep="\t")
  write.table(pw_mat,file=paste(prefix,"AverageProximity",sample,"Pairwise","txt",sep="."), sep="\t")

}

```


### contruct neighborhood matrix ######################################################
```{r}

Samplelist = unique(data.combined@meta.data$ID)
Imagelist = Images(object = data.combined)

for(i in 1:length(Samplelist)){
  image = Imagelist[i]
  sample = Samplelist[i]
  
  table_coord <- data.combined[[image]]@coordinates
  table_coord <- table_coord[order(table_coord$row, table_coord$col),]

  neighbor_table <- data.frame(matrix(nrow=nrow(table_coord),ncol=nrow(table_coord)))
  rownames(neighbor_table) = rownames(table_coord)
  colnames(neighbor_table) = rownames(table_coord)

  neighbor_table[is.na(neighbor_table)]<-0
  
  for(j in 1:nrow(table_coord)){
    for(k in 1:nrow(table_coord)){
      
      row_j = table_coord[j,]$row
      col_j = table_coord[j,]$col
      
      row_k = table_coord[k,]$row
      col_k = table_coord[k,]$col

      # if(row_k == row_j & col_k == col_j){
      #   neighbor_table[j,k] = 1
      # }
      
      if(row_k == row_j & col_k == col_j-2){
        neighbor_table[j,k] = 1
      }
      
      if(row_k == row_j & col_k == col_j+2){
        neighbor_table[j,k] = 1
      }
      
      if(row_k == row_j-1 & col_k == col_j-1){
        neighbor_table[j,k] = 1
      }
      
      if(row_k == row_j-1 & col_k == col_j+1){
        neighbor_table[j,k] = 1
      }
      
      if(row_k == row_j+1 & col_k == col_j-1){
        neighbor_table[j,k] = 1
      }
      
      if(row_k == row_j+1 & col_k ==col_j+1){
        neighbor_table[j,k] = 1
      }      
    }
  }

  write.table(neighbor_table,file=paste(prefix,"NeighborMatrix",sample,"txt",sep="."), sep="\t")
}

```


### calculate the direct neighboring fraction of each well ########
```{r}

Samplelist = unique(data.combined@meta.data$ID)
Imagelist = Images(object = data.combined)
AUC_Binary <- read.delim(file=paste(prefix,"AUC_Binary","txt",sep="."),sep="\t",header=T,row.names=1)

neighbor_prop <- data.frame(matrix(nrow=0,ncol=ncol(AUC_Binary)))
colnames(neighbor_prop) = colnames(AUC_Binary)

for(i in 1:length(Samplelist)){
  image = Imagelist[i]
  sample = Samplelist[i]
  
  neighbor_table=read.delim(paste("Visium.MetsGBM.CombinedAnalysis.Aug2022.AUCell.MetsGBMscRNAseqCelltype","NeighborMatrix",sample,"txt",sep="."),sep="\t",header=T,row.names=1, check.names=FALSE)
  
  #neighbor_table[4,neighbor_table[4,]==1]
  
  #subset = AUC_Binary[AUC_Binary[,anchor_type]==1 & rownames(AUC_Binary) %in% rownames(neighbor_table) & !is.na(AUC_Binary[,anchor_type]),]
  #subset = AUC_Binary[rownames(AUC_Binary) %in% rownames(neighbor_table) & !is.na(AUC_Binary[,anchor_type]),]
  subset = AUC_Binary[rownames(AUC_Binary) %in% rownames(neighbor_table) & rowSums(is.na(AUC_Binary))<=0,]


  anchor_wells = rownames(subset)
  
  for(j in 1:length(anchor_wells)){
    ids = rownames(neighbor_table[neighbor_table[anchor_wells[j],]==1,])
    AUC_sub = AUC_Binary[c(anchor_wells[j],ids),]
    
    prop <- apply(AUC_sub, 2,function(x,y) (sum(x)/y),  nrow(AUC_sub))
    neighbor_prop[anchor_wells[j],] <- prop
  }
}

write.table(neighbor_prop,file=paste(prefix,"NeighborProp","txt",sep="."), sep="\t")
#neighbor_prop <- read.delim(file=paste(prefix,"NeighborProp","txt",sep="."),sep="\t",header=T,row.names=1)

# plot the neighborhood fraction of selected anchor wells ######
anchor_type = "CD8.CXCL13.LAG3.Tex"
subset_ids = rownames(AUC_Binary[AUC_Binary[,anchor_type]==1 & !is.na(AUC_Binary[,anchor_type]),])

data.sub = subset(data.combined, cells = subset_ids)
neighbor_prop_sub = neighbor_prop[colnames(data.sub),]

data.sub@assays[['NeighborProp']] <- CreateAssayObject(data = t(neighbor_prop_sub))
data.sub@assays$NeighborProp@key <- "newkey_"

DefaultAssay(data.sub) <- "NeighborProp"

types = colnames(neighbor_prop_sub)

outfile = paste(prefix,"NeighborProp",anchor_type,"All","png",sep=".")
png(filename = outfile,width = 1500, height = 10000)
SpatialFeaturePlot(data.sub, features = types)
dev.off()

data = neighbor_prop_sub
data$ID = data.sub@meta.data[,"ID"]
data$ID = factor(data$ID, levels =                      c("SSW-18-16692_A1_M","SSW-19-15481_A1_M_P","SSW-18-22490_A1_L","SSW-19-07361_B2_L_P","SSW-19-19183_A2","SSW-19-02982_K2_P"))

colors =c("pink","red","#FAD5A5","orange","cyan","blue")

ncol = 3
nrow1 = length(types)/ncol
height1 = nrow1*480
width1 = length(levels(data$ID))*180*ncol

outfile = paste(prefix,"NeighborProp",anchor_type,"All","Boxplot","png",sep=".")
png(filename = outfile,width = width1, height = height1)
panels=list()
l=length(colnames(data))
i=1
for(i in 1:length(types)){
#for(i in 1:2){
  celltype = types[i]
  data2 = data[,c(i,l)]
  colnames(data2)=c("normExpr","ID")

  y = ggplot(data=data2, aes(x=ID,y=normExpr,fill=ID)) +
       geom_violin(trim = FALSE)+ylim(0,1)+
        #stat_summary(fun="median",geom="crossbar",width=0.3,color="black")+
        #geom_jitter(width=0.3, alpha=0.3) +
       geom_boxplot(outlier.colour="black",notch=TRUE,outlier.size=5)+
        scale_fill_manual(values=colors)+
      theme(
        axis.text.x = element_blank(),
        axis.text.y = element_text(size=20),
        axis.ticks.x = element_blank(),
        axis.title = element_blank(),
        legend.position = "none",
        plot.title=element_text(face = "italic",size=20))+
      ggtitle(celltype)
    
      panels[[i]]=y
  
}
ggarrange(plotlist=panels,ncol=ncol,nrow=nrow1)
dev.off()


outfile = paste(prefix,anchor_type,"NeighborProp","WilcoxTest","txt",sep=".")

for(i in 1:ncol(neighbor_prop_sub)){
  celltype = colnames(neighbor_prop_sub)[i]
  data = data.frame(prop = as.numeric(neighbor_prop_sub[,i]),ID = data.sub@meta.data[,"ID"])
  
  data$ID = factor(data$ID, levels =
                     c("SSW-18-16692_A1_M","SSW-19-15481_A1_M_P","SSW-18-22490_A1_L","SSW-19-07361_B2_L_P","SSW-19-19183_A2","SSW-19-02982_K2_P"))
  
  sub1 <- subset(data,ID=="SSW-18-16692_A1_M")
  sub2 <- subset(data,ID=="SSW-19-15481_A1_M_P")
  t <- wilcox.test(sub1$prop,sub2$prop,paired=F)
  p = t$p.value
  #delta = median(sub2$prop)-median(sub1$prop)
  line = paste(celltype, "MelanomavsMelanoma.pembro", p, sep = "\t")
  write(line,file=outfile,append=TRUE)
  
  sub1 <- subset(data,ID=="SSW-18-22490_A1_L")
  sub2 <- subset(data,ID=="SSW-19-07361_B2_L_P")
  t <- wilcox.test(sub1$prop,sub2$prop,paired=F)
  p = t$p.value
  line = paste(celltype, "LungvsLung.pembro", p, sep = "\t")
  write(line,file=outfile,append=TRUE)  

  sub1 <- subset(data,ID=="SSW-19-19183_A2")
  sub2 <- subset(data,ID=="SSW-19-02982_K2_P")
  t <- wilcox.test(sub1$prop,sub2$prop,paired=F)
  p = t$p.value
  line = paste(celltype, "GBMvsGBM.pembro", p, sep = "\t")
  write(line,file=outfile,append=TRUE)  
}

```


### calculate the number/fraction of tumor-adjacent cell type ############
```{r}

Samplelist = unique(data.combined@meta.data$ID)
AUC_Binary <- read.delim(file=paste(prefix,"AUC_Binary","txt",sep="."),sep="\t",header=T,row.names=1)
types = colnames(AUC_Binary)

prop_cut = 0.5

adjacent_type = "Tumor"
types <- types[!types %in% c(adjacent_type)]
#types <- types[!types %in% c("Vascular","BrainNormal")] #only check immune cell subtypes


neighbor_num <- data.frame()

for(i in 1:length(Samplelist)){
  sample = Samplelist[i]
    
  neighbor_table=read.delim(paste("Visium.MetsGBM.CombinedAnalysis.Aug2022.AUCell.MetsGBMscRNAseqCelltype","NeighborMatrix",sample,"txt",sep="."),sep="\t",header=T,row.names=1, check.names=FALSE)
  
  id_total = c()
  
  for(j in 1:length(types)){
    type = types[j]
  
    subset = AUC_Binary[rownames(AUC_Binary) %in% rownames(neighbor_table) & !is.na(AUC_Binary[,type]) & AUC_Binary[,type]==1,]
    
    if(nrow(subset)==0){
      neighbor_num[i,type] = 0
      id_total = id_total
      next
    } 
      
    anchor_wells = rownames(subset)
    
    n_adjacent = 0
    id_type = c()

    for(k in 1:length(anchor_wells)){
      ids = rownames(neighbor_table[neighbor_table[anchor_wells[k],]==1,])
      AUC_sub = AUC_Binary[c(anchor_wells[k],ids),adjacent_type]
      
      AUC_sub[is.na(AUC_sub)] <- 0
    
      prop <- sum(AUC_sub)/length(AUC_sub)
      if(prop > prop_cut){
        n_adjacent = n_adjacent+1
        id_type <- c(id_type,anchor_wells[k])
      }
    }
    neighbor_num[i,type] = n_adjacent
    id_total = union(id_total, id_type)
  }
 
  neighbor_num[i,"TotalAdjacent"] = length(id_total) 
  
}

rownames(neighbor_num) = Samplelist


#calculate celltype adjacent total T and myeloid cells#############
###################################################################
types_Tcell = c("CD4.CXCL13.CTLA4.Tex","CD4.IL7R.CCR7.CD40LG.Tcm",
                "CD8.CXCL13.LAG3.Tex","CD8.GZMH.GZMA.CD52.Teff",
                "CD8.GZMK.CCL4.CCL5.Teff","CD8.IL7R.CD69.Tm",
                "CD8.ISG.High","CD8.ITGA1.ITGAL.Trm",
                "CD8.TCF7.CD226.Tprog.exh","CD8.VCAM1.IFNG.Tex",
                "CTL.NK","CTL.NK.XCL1.XCL2",
                "IL7R.KLRB1.Tm","Treg")


types_Myeloid = c("cDC1","cDC2",
                  "Complement.RiboHigh","Microglia",
                  "Microglia.ISG.IL10","Microglia.Phagocytosis",
                  "Monocyte","Monocyte.ISG",
                  "Mphage.Angiogenesis","Mphage.IL1B.CCL3.CCL4",
                  "Mphage.ISG.MRC1.LILRB2","Mphage.Lysosome",
                  "Mphage.MRC1.LYVE1","pDC")


n_Tcell = n_myeloid = c()
for(i in 1:length(Samplelist)){
  sample = Samplelist[i]
    
  neighbor_table=read.delim(paste("Visium.MetsGBM.CombinedAnalysis.Aug2022.AUCell.MetsGBMscRNAseqCelltype","NeighborMatrix",sample,"txt",sep="."),sep="\t",header=T,row.names=1, check.names=FALSE)
  
  subset = AUC_Binary[rownames(AUC_Binary) %in% rownames(neighbor_table),types_Tcell]
  subset <- subset[rowSums(subset)>=1,]
    
  anchor_wells = rownames(subset)
    
  n_adjacent = 0

  for(k in 1:length(anchor_wells)){
    ids = rownames(neighbor_table[neighbor_table[anchor_wells[k],]==1,])
    AUC_sub = AUC_Binary[c(anchor_wells[k],ids),adjacent_type]
      
    AUC_sub[is.na(AUC_sub)] <- 0
    
    prop <- sum(AUC_sub)/length(AUC_sub)
    if(prop > prop_cut){
      n_adjacent = n_adjacent+1
      #id_type <- c(id_type,anchor_wells[k])
    }
  }
  
  
  n_Tcell = c(n_Tcell, n_adjacent)

  subset = AUC_Binary[rownames(AUC_Binary) %in% rownames(neighbor_table),types_Myeloid]
  subset <- subset[rowSums(subset)>=1,]
    
  anchor_wells = rownames(subset)
    
  n_adjacent = 0

  for(k in 1:length(anchor_wells)){
    ids = rownames(neighbor_table[neighbor_table[anchor_wells[k],]==1,])
    AUC_sub = AUC_Binary[c(anchor_wells[k],ids),adjacent_type]
      
    AUC_sub[is.na(AUC_sub)] <- 0
    
    prop <- sum(AUC_sub)/length(AUC_sub)
    if(prop > prop_cut){
      n_adjacent = n_adjacent+1
      #id_type <- c(id_type,anchor_wells[k])
    }
  }
  n_myeloid = c(n_myeloid, n_adjacent)

}

neighbor_num$AdjacentTcell = n_Tcell
neighbor_num$AdjacentMyeloid = n_myeloid

write.table(neighbor_num,file=paste(prefix,"CellNumAdjacentTo",adjacent_type,"txt",sep="."), sep="\t")


```



### calculate the direct neighboring fraction of each well of a particular anchor type ############
```{r}
neighbor_prop <- read.delim(file=paste(prefix,"NeighborProp","txt",sep="."),sep="\t",header=T,row.names=1)

#anchor_types = c("Mphage.MRC1.LYVE1","Mphage.Angiogenesis","Mphage.Lysosome")
anchor_types = c("CD8.CXCL13.LAG3.Tex","CD8.VCAM1.IFNG.Tex")


# plot the neighborhood fraction of selected anchor wells ######
#subset_ids = rownames(AUC_Binary[AUC_Binary[,anchor_types]==1 & !is.na(AUC_Binary[,anchor_type]),])
subset_ids = rownames(AUC_Binary[rowSums(AUC_Binary[,anchor_types])>=1 & rowSums(is.na(AUC_Binary))<=0,])

data.sub = subset(data.combined, cells = subset_ids)

neighbor_prop.sub = neighbor_prop[colnames(data.sub),]
neighbor_prop.sub$ID = data.sub@meta.data[,"ID"]

##output anchor wells adjacent to specific cell subtypes ##
samples_sub = c("SSW-18-16692_A1_M","SSW-19-15481_A1_M_P","SSW-18-22490_A1_L","SSW-19-07361_B2_L_P")

subset <- neighbor_prop.sub[neighbor_prop.sub$Tumor<0.5 & neighbor_prop.sub$Vascular>0.5 & neighbor_prop.sub$ID %in% samples_sub,]
subset <- na.omit(subset)
ids = rownames(subset)
fileConn<-file(paste(prefix,"CD8.Tex","AdjacentTo","Vascular","txt",sep="."))
writeLines(ids, fileConn)

subset <- neighbor_prop.sub[neighbor_prop.sub$Tumor>0.5 & neighbor_prop.sub$Vascular<0.5 & neighbor_prop.sub$ID %in% samples_sub,]
subset <- na.omit(subset)
ids = rownames(subset)
fileConn<-file(paste(prefix,"Mphage.MRC1","AdjacentTo","Tumor","txt",sep="."))
writeLines(ids, fileConn)




```




