```{r}
library(Seurat)
library(Matrix)
library(dplyr)
library(cowplot)
require(data.table)
library(harmony)
library("ggplot2")

options(bitmapType='cairo') 
prefix = "Seurat.MetsGBM.CombinedAnalysis.Apr2022"

infile1 = "scRNAseq/GBM2021.rec.TIL.Normalized.VarGene.rds"
infile2 = "scRNAseq/Mets2022.TIL.Normalized.VarGene.rds"
infile3 = "scRNAseq/WerbCell2022.ImmuneCombined.Apr2022.Normalized.VarGene.rds"
infile4 = "scRNAseq/Feldman2018.Combined.Apr2022.BrainMets.Normalized.VarGene.rds"


```

#read in rds files and create seurat object lists for integration
```{r}
data.gbm.own = readRDS(file=infile1)
data.mets.own = readRDS(file=infile2)
data.mets.werb = readRDS(file=infile3)
data.mets.Feldman = readRDS(file=infile4)

data.gbm.own$Dataset <- "GBM.2021"
data.mets.own$Dataset <- "Mets.2022"
data.mets.werb$Dataset <- "Werb.2022"
data.mets.Feldman$Dataset <- "Feldman"


#create multiple seurat objects depending on their tumor type and treatment group
#this way the most variable genes we selected during integration should be evenly distributed across tumor types and treatment groups
ob.list<-SplitObject(data.gbm.own,split.by= "condition")

ob.list.gbm.rec <- SplitObject(ob.list$GBM.rec,split.by= "ID")
gbm.merge <- merge(
  x = ob.list.gbm.rec[[1]],
  y = unlist(ob.list.gbm.rec[2:length(ob.list.gbm.rec)])
)
gbm.merge@assays$RNA@var.features <- SelectIntegrationFeatures(ob.list.gbm.rec)


ob.list.gbm.pembro <- SplitObject(ob.list$GBM.pembro,split.by= "ID")
gbm.pembro.merge <- merge(
  x = ob.list.gbm.pembro[[1]],
  y = unlist(ob.list.gbm.pembro[2:length(ob.list.gbm.pembro)])
)
gbm.pembro.merge@assays$RNA@var.features <- SelectIntegrationFeatures(ob.list.gbm.pembro)



ob.list<-SplitObject(data.mets.own,split.by= "condition")
ob.list.met1 <- SplitObject(ob.list$met,split.by= "ID")
ob.list.met.pembro1 <- SplitObject(ob.list$met.pembro,split.by= "ID")


ob.list.met2 <- SplitObject(data.mets.werb,split.by= "ID")

ob.list.met <- c(unlist(ob.list.met1),unlist(ob.list.met2))
met.merge <- merge(
  x = ob.list.met[[1]],
  y = unlist(ob.list.met[2:length(ob.list.met)])
)
met.merge@assays$RNA@var.features <- SelectIntegrationFeatures(ob.list.met)


ob.list.met.pembro <- c(unlist(ob.list.met.pembro1),data.mets.Feldman)
met.pembro.merge <- merge(
  x = ob.list.met.pembro[[1]],
  y = unlist(ob.list.met.pembro[2:length(ob.list.met.pembro)])
)
met.pembro.merge@assays$RNA@var.features <- SelectIntegrationFeatures(ob.list.met.pembro)


ob.list = list(gbm.merge, gbm.pembro.merge, met.merge, met.pembro.merge)

rm(data.gbm.own)
rm(data.mets.own)
rm(data.mets.Feldman)
rm(data.mets.werb)
rm(ob.list.met.pembro1)
rm(ob.list.met.pembro)
rm(ob.list.met)
rm(ob.list.met1)
rm(ob.list.met2)
rm(ob.list.gbm.rec)
rm(ob.list.gbm.pembro)
rm(gbm.merge)
rm(gbm.pembro.merge)
rm(met.merge)
rm(met.pembro.merge)

gc()

```

#Run integration, scaling and PCA
```{r}
#ob.list = c(ob.list.own, ob.list.public)

k.filterMax <- min(200, min(sapply(ob.list, ncol)))
anchors <- FindIntegrationAnchors(object.list = ob.list, k.filter = k.filterMax, anchor.features = 2000)
data.combined <- IntegrateData(anchorset = anchors, k.weight = k.filterMax)

saveRDS(anchors, file=paste(prefix, "anchors", "rds", sep="."))
rm(anchors)

DefaultAssay(object = data.combined) <- "integrated"

data.combined <- ScaleData(object = data.combined, verbose = TRUE, vars.to.regress = c("percent.mt","percent.ribo","nFeature_RNA","nCount_RNA","S.Score","G2M.Score"))

gc()

data.combined <- RunPCA(object = data.combined, npcs = 30, verbose = TRUE)


saveRDS(data.combined, file=paste(prefix, "rds", sep="."))

gc()
```

#Run harmony
```{r}

data.combined <- RunHarmony(data.combined, group.by.vars="Dataset",assay.use = "integrated", plot_convergence = TRUE)
harmony_embeddings <- Embeddings(data.combined, 'harmony')

gc()
```


#Run UMAP for cluster visualization
```{r}
minDist = 0.1
spread1 = 5
NN = 20
seedUse = 122
epoch = 1000

prefix1 = paste(prefix,".MinD",minDist,".sprd",spread1,".NN",NN,".seed",seedUse,".ep",epoch,sep="")

data.combined <- RunUMAP(object = data.combined, reduction = "harmony", dims = 1:30,min.dist = minDist, spread=spread1, n.neighbors=NN, seed.use = seedUse, n.epochs=epoch)

outfile = paste(prefix1,"_UMAP", ".nUMI.png",sep="")
png(filename = outfile,width = 800, height = 800)
FeaturePlot(object = data.combined, features = c("nCount_RNA"), min.cutoff = "q9", pt.size = 2) #umi plot
dev.off()

outfile = paste(prefix1,"_UMAP",".nGene.png",sep="")
png(filename = outfile,width = 800, height = 800)
FeaturePlot(object = data.combined, features = c("nFeature_RNA"), min.cutoff = "q9", pt.size = 2) #umi plot
dev.off()

outfile = paste(prefix1,"_UMAP",".pctRibo.png",sep="")
png(filename = outfile,width = 800, height = 800)
FeaturePlot(object = data.combined, features = c("percent.ribo"), min.cutoff = "q9", pt.size = 2) #umi plot
dev.off()

outfile = paste(prefix1,"_UMAP",".pctMito.png",sep="")
png(filename = outfile,width = 800, height = 800)
FeaturePlot(object = data.combined, features = c("percent.mt"), min.cutoff = "q9", pt.size = 2) #umi plot
dev.off()

data.combined@meta.data$Dataset = factor(data.combined@meta.data$Dataset, levels = 
                                          c("GBM.2021","Mets.2022","Werb.2022","Feldman"))


outfile = paste(prefix1,"_UMAP.dataset.png",sep="")
png(filename = outfile,width = 800, height = 800)
DimPlot(object = data.combined, reduction = "umap", group.by = "Dataset", label = TRUE, label.size = 8, pt.size = 0.5)
dev.off()

outfile = paste(prefix1,"_UMAP.dataset.split.png",sep="")
png(filename = outfile,width = 1600, height = 800)
DimPlot(object = data.combined, reduction = "umap", group.by = "Dataset", split.by = "Dataset", label = FALSE, label.size = 8, pt.size = 0.5)
dev.off()


data.combined@meta.data$condition = factor(data.combined@meta.data$condition, levels = 
                                          c("GBM.rec","GBM.pembro","met","met.pembro"))
color_condition <- c("cyan","blue","pink","red")

outfile = paste(prefix1,"_UMAP.condition.split.png",sep="")
png(filename = outfile,width = 1600, height = 800)
DimPlot(object = data.combined, reduction = "umap", group.by = "condition", split.by = "condition", label = TRUE, label.size = 8, pt.size = 0.5, cols = color_condition)
dev.off()

outfile = paste(prefix1,"_UMAP.WerbCelltype.png",sep="")
png(filename = outfile,width = 800, height = 800)
DimPlot(object = data.combined, reduction = "umap", group.by = "Cell_Type", label = TRUE, label.size = 8, pt.size = 1, na.value = "lightgrey", raster=FALSE)
dev.off()

saveRDS(data.combined, file=paste(prefix1, "rds", sep="."))


```


#Clustering
```{r}
DefaultAssay(object = data.combined) <- "integrated"

data.combined <- FindNeighbors(object = data.combined, reduction = "harmony", dims = 1:30, nn.method = "annoy", annoy.metric = "cosine", force.recalc = TRUE)
data.combined <- FindClusters(data.combined, resolution = 0.7)

#colors = c("red","orange","yellow","green","lightgreen","blue","magenta","cyan","pink","brown","black","darkblue","darkgrey","lightgrey","maroon","lightblue")

outfile = paste(prefix1,"_UMAP", ".Res0.7.png",sep="")
png(filename = outfile,width = 800, height = 800)
DimPlot(object = data.combined, reduction = "umap", group.by = "integrated_snn_res.0.5", label = TRUE, label.size = 8, pt.size = 1)#, cols = colors[1:8])
dev.off()

saveRDS(data.combined, file=paste(prefix1, "rds", sep="."))
#data.combined = readRDS(file=paste(prefix1, ".rds", sep=""))

```


#Find cluster markers
```{r}
DefaultAssay(data.combined) <- "RNA"
Idents(object = data.combined) <- "integrated_snn_res.0.3"
cluster.markers<- FindAllMarkers(object = data.combined, only.pos = TRUE, min.pct = 0.25,logfc.threshold = 0.322,test.use="MAST")
outfile = paste(prefix1,"_ClusterDiffGene.Res0.3", ".csv",sep="")
write.csv(cluster.markers, file=outfile)

saveRDS(data.combined,file=paste(prefix1, ".rds", sep=""))
#data.combined = readRDS(file=paste(prefix1, ".rds", sep=""))

```

#Find cluster markers
```{r}

DefaultAssay(data.combined) <- "RNA"

outfile = paste(prefix,"_UMAP.", "MainCellTypes.png",sep="")
png(filename = outfile,width = 2000, height = 2000)
FeaturePlot(object = data.combined, features = c("EPCAM","PECAM1","PDGFRB","GFAP","OLIG2","PTPRC","CD3D","CD8B","CD4","CD14","FCGR3A","S100A9","C1QA","HLA-DRB1","MRC1","CD163","LYVE1","CD79A","TMEM119","P2RY12","CX3CR1","MLANA","PMEL","EGFR","MKI67"), min.cutoff = "q9",  pt.size = 1, ncol=5, reduction = "umap", cols = c("lightgrey","red"))
dev.off()

outfile = paste(prefix,"_UMAP.MainCellTypes2.png",sep="")
png(filename = outfile,width = 1200, height = 1500)
FeaturePlot(object = data.combined, features = c("CD3D","CD14","HLA-DRA","PTPRC","S100A8","APOE","CLEC10A","SOX2","GFAP","P2RY12","OLIG1","PTPRZ1","SOX4","CD24","PMEL","MLANA","EPCAM","MKI67"), min.cutoff = "q9",  pt.size = 1, ncol=4, cols = c("lightgrey","red"))
dev.off()

outfile = paste(prefix,"_UMAP.MainCellTypes3.png",sep="")
png(filename = outfile,width = 1200, height = 1500)
FeaturePlot(object = data.combined, features = c("PTPRC","CD3D","CD8B","CD4","CD14","CD79A","HLA-DRA","P2RY12","GFAP","OLIG1","PTPRZ1","PMEL","MLANA","CLDN5","PECAM1","RGS5","ACTA2","KRT19","EPCAM","MKI67"), min.cutoff = "q9",  pt.size = 1, ncol=4, cols = c("lightgrey","red"))
dev.off()



```


#assign cell types ###
```{r}
#assign each cluster cell type
Idents(object = data.combined) <- "integrated_snn_res.0.3"
current.cluster.ids <- c(0:18)
new.cluster.ids <- c("Lymphoid","Myeloid","Myeloid",#2
                     "Ambiguous","Tumor/normal","Myeloid",#5
                     "Tumor/normal","Myeloid","Myeloid",#8
                     "Ambiguous","Tumor/normal","Lymphoid.Cycling",#11
                     "Myeloid.Cycling","Tumor/normal","Tumor/normal",#14
                     "Tumor/normal","Tumor/normal","Tumor/normal",#17
                     "Tumor/normal") #18


Idents(object = data.combined) <- plyr::mapvalues(x = Idents(object = data.combined), from = current.cluster.ids, to = new.cluster.ids)

data.combined@meta.data$celltype <- Idents(object = data.combined)
data.combined@meta.data$celltype <- factor(data.combined@meta.data$celltype, levels = c("Lymphoid","Myeloid","Lymphoid.Cycling","Myeloid.Cycling","Tumor/normal","Ambiguous"))

library(colorspace)
library(scales)
colorsCluster = c(qualitative_hcl(5),"grey")

outfile = paste(prefix1,"_UMAP.Celltype.png",sep="")
png(filename = outfile,width = 800, height = 800)
plot <- DimPlot(object = data.combined, reduction = "umap", group.by = "celltype", label = TRUE, label.size = 8, pt.size = 1, cols = colorsCluster, raster=FALSE)
dev.off()

#outfile = paste(prefix1,"_UMAP.Celltype.noLabel.png",sep="")
#png(filename = outfile,width = 800, height = 800)
outfile = paste(prefix1,"_UMAP.Celltype.noLabel.tiff",sep="")
tiff(file=outfile, width=3500, height=3000, res=600)
DimPlot(object = data.combined, reduction = "umap", group.by = "celltype", label = FALSE, label.size = 20, pt.size = 1, cols = colorsCluster, raster=FALSE)
dev.off()

saveRDS(data.combined,file=paste(prefix1, ".rds", sep=""))
#data.combined = readRDS(file=paste(prefix1, ".rds", sep=""))
```


## Write the files into separate populations for subset analysis
```{r}
ids = rownames(subset(data.combined@meta.data,celltype %in% c("Lymphoid","Lymphoid.Cycling")))
fileConn<-file(paste(prefix1,".Lymphoid.ID.txt",sep=""))
writeLines(ids, fileConn)

ids = rownames(subset(data.combined@meta.data,celltype %in% c("Myeloid","Myeloid.Cycling")))
fileConn<-file(paste(prefix1,".Myeloid.ID.txt",sep=""))
writeLines(ids, fileConn)

ids = rownames(subset(data.combined@meta.data,celltype %in% c("Tumor/normal","Ambiguous")))
fileConn<-file(paste(prefix1,".NonImmune.Ambiguous.txt",sep=""))
writeLines(ids, fileConn)

```



#calculate each cell type's fraction in each sample ###
```{r}
## Enumerate per sample
totalNum=c(table(data.combined@meta.data$ID)[[1]])
Group = names(table(data.combined@meta.data$ID))
table_sub = data.frame()
table_sub = subset(data.combined@meta.data, ID==Group[1])
percentages=as.data.frame((table(table_sub$celltype)))$Freq
table1 = percentages

m1 = data.combined@meta.data

for(i in 2:length(Group)){
  id = Group[i]
  n_total = table(data.combined@meta.data$ID)[[i]]
  totalNum=c(totalNum,n_total)
  
  table_sub = data.frame()
  table_sub = subset(data.combined@meta.data, ID==id)
  
  percentages=as.data.frame((table(table_sub$celltype)))$Freq
  table1=as.data.frame(rbind(table1,percentages))
}

table2 = as.data.frame(cbind(table1,Total=totalNum))
rownames(table2)=Group
colnames(table2)=c(levels(data.combined@meta.data$celltype),"Total")
table2$CD45 = table2$Lymphoid+table2$Myeloid+table2$Lymphoid.Cycling+table2$Myeloid.Cycling

sampleMap = unique(data.combined@meta.data[,c("ID","condition","CancerType")])
rownames(sampleMap) = sampleMap$ID
sampleMap = sampleMap[Group,]
table2$condition = sampleMap$condition
table2$CancerType = sampleMap$CancerType

outfile = paste(prefix,"_CelltypeByID.txt",sep="")
write.table(table2,file=outfile,sep="\t")

```


#generate bargraph or boxplots of different populations
```{r}
library(dplyr)
library(ggplot2)
library(reshape)

##use table2 to generate percentages value
table3 = table2
#table3$Lymphoid = table2$`Lymphoid`+table2$`Lymphoid.Cycling`
#table3$Myeloid = table2$`Myeloid`+table2$`Myeloid.Cycling`


## compute the fraction against total
table3 = table3 %>% rowwise()
tableCD45 = table3 %>%  mutate_if(~ is.numeric(.x), ~ .x/CD45)
tableCD45 = as.data.frame(tableCD45[,c(1:4,9)])
rownames(tableCD45) = rownames(table2)
tableCD45$ID = rownames(tableCD45)

tableCD45_melt = melt(tableCD45,id.vars=c("ID","condition"))
colnames(tableCD45_melt) = c("ID","Condition","Celltype","Fraction")
colors = c("cyan","blue","pink","red")

tableCD45_melt$Condition = factor(tableCD45_melt$Condition, levels = c("GBM.rec","GBM.pembro","met","met.pembro"))
tableCD45_melt$Celltype = factor(tableCD45_melt$Celltype, levels = c("Lymphoid","Myeloid","Lymphoid.Cycling","Myeloid.Cycling"))

width1 = length(levels(tableCD45_melt$Celltype))*200
outfile = paste(prefix,"CellFractionBoxplot.AllCell.png",sep="_")
png(filename = outfile,width = width1, height = 600)
y = ggplot(data=tableCD45_melt, aes(x=Celltype,y=Fraction,fill=Condition)) +
       #geom_violin(trim = FALSE)+
       geom_boxplot(outlier.colour="black",width=0.8,outlier.shape=1,outlier.size=0)+
       geom_point(aes(fill = Condition), size = 2, shape = 21, position = position_jitterdodge())+
        scale_fill_manual(values=rep(colors,length(levels(tableCD45_melt$Celltype))))+
      theme(
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(size=20),
        axis.text.y = element_text(size=28),
        axis.ticks.x = element_blank(),
        axis.title = element_blank(),
        legend.position = "none",
        plot.title=element_text(face = "italic",size=32))+
      ggtitle("CD45 cell fraction")
y
dev.off()
```

## get the wilcoxon p-value
```{r}
celltypeL = c()
metsPval=c()
GBMPval=c()

for(i in 1:4){
  celltypeL = c(celltypeL,colnames(tableCD45)[i])
  a = wilcox.test(tableCD45[[i]]~tableCD45$condition,subset=tableCD45$condition %in% c("GBM.rec","GBM.pembro"))
  GBMPval = c(GBMPval,a$p.value)
  
  b = wilcox.test(tableCD45[[i]]~tableCD45$condition,subset=tableCD45$condition %in% c("met","met.pembro"))
  metsPval = c(metsPval,b$p.value)
}

table1 = data.frame(Type=celltypeL,mets.vs.Pembro.wilcox=metsPval,GBM.vs.Pembro.wilcox=GBMPval)
outfile = paste(prefix,"_CelltypeFracDiff.WilcoxPval.csv",sep="")
write.csv(table1,file=outfile)
```

